---
title: "FinalAssignment2"
author: "Emily Ferry"
date: "1/27/2020"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(gapminder)
library(ggthemes)
library(dplyr)
library(tidyr)
library(ISLR)
```

##Introduction 

I started as I normally do, trying to find a dataset that interests me in my field. I find it hard to find marketing data that makes sense as most quality data needs to be paid for. The sets I do find usually have some sort of questionable background, leaving me unsure of what to be asking about them. For instance, I saw a set (several really) on kaggle that was about banks marketing- however this was done in protugue and some of the variables were hard to wrap my head around. I also am not able to access work data quite yet. I did make attempts to use an Avocado dataset from Kaggle, however I didnt really have a question/ interest in this data and there seemed to be not ony a massive amount of varaibles to try and plot, but also some gaps in the data. Next  I attempted to use R datasets such as OJ and USArrests, again though they were not working out. 

The data I settled on was R's Longley's Economic Regression Data. While is is a macroeconomc data set that is well know for a highly collinear regression, I needed to settle on something and this ecomomic dataset has some interesting variables, such as GNP rather then GDP and number of people in the armed forces. The data was not very timley as it was from 1947 to 1962, but again as I am not big on ecomonics it did interest me in the slightest and I thought it would result in some nice visuals to help me with my questions. I was cuirious if  an increased number of armed forces affect the GNP? how is GNP affected over time by populated and those unemployed vs. employed. I very likely already know these answers do to the high collinear regression, however it is still interesting to explore how these factors might affect a countries overall economic standings. 


#The Data
```{r}
longley
?longley #A macroeconomic data set which provides a well-known example for a highly collinear regression.A data frame with 7 economical variables, observed yearly from 1947 to 1962 (n=16).

```

GNP is the GNP (Gross National Product) is a measure of national income.  It is the market value of all goods and services created in a given year by labor and property that belong to the citizens of a given country. It measures how much the goods and services that are produced cost.  

#some visuals 

```{r}
ggplot(data = longley,aes(x = Population, y = Unemployed))+
  geom_point()

ggplot(longley, aes(x=Year, y=Employed)) +
  geom_point() + stat_smooth(method="lm") #adding a loess line with a linear regression model 


ggplot(longley, aes(x=Year, y=Employed, color = GNP)) +
  geom_point()  # looking at # of those employeed over time ( year ) showing the GNP increase visually in color 

ggplot(longley, aes(x=Year, y=Armed.Forces, color = GNP)) +
  geom_point()  # same as the last but with armed forces # it looks like maybe there is a gap in the data or something else is occuring


ggplot(data = longley, aes(x= GNP, y = Employed, color = Armed.Forces, size = Population)) +
  geom_point() + 
  stat_smooth( method = lm) # this is sort of interesting, a liner regression model with a loess line showing the employement number by GNP, we also see the size of armed forces prevenalt with the color variation, dark being less light being more, but it gets a little intese with population also being reflected in the size of the data points, the bigger the point the larger the population 

ggplot(data = longley, aes(x= Year, y = Employed, size = Armed.Forces)) +
  geom_point() + 
  stat_smooth( method = lm) # I wonder if those employed are also counted as those in Armed.Forces 


ggplot(data = longley, aes(x= Year, y = GNP.deflator, size = Armed.Forces)) +
  geom_point()  # I wonder if those employed are also counted as those in Armed.Forces, in 1947-53 we have the Cold War, in the 60's there was the cuban missle crisis and the vietnam war 
ggplot(data = longley, mapping=  aes(x= GNP.deflator, y = Employed)) +
  geom_point( alpha = .4) + 
  stat_smooth( method = lm) + 
  facet_wrap(~Population) # that was useless 

```

# In Conclusion : 

```{r}

# Install
install.packages("wesanderson")
# Load
library(wesanderson)


p2<-ggplot(data = longley, aes(x= Year, y = GNP.deflator, color = Armed.Forces, size = Population)) +
  geom_point() + 
  stat_smooth( method = lm)+
  xlab("Year")+
  ylab("GNP Adjusted for Inflation") +
  ggtitle("GNP by Year", subtitle = "Population by Size and Number of Armed Forces by Gradient")+
  theme_minimal()

p2+scale_color_gradient(low="#000099", high="#6699FF")

?longley


```

This graph is maybe a little confusing at first, and takes a few minutes to figure out, if I were presenting this I would definitly have made soemthing a little simplier (only one type of scale versus two) I think I liked this becuase it helped me to awnser some questions I originally had and it even brought one to the forefront of my data. Even though I knew the data was a highly correlated regression, I was able to visually see that as as the over the years GNP.deflator ( adjsuted GNP ) increased, as did the population while you can see a spike in the number of armed forced increased around 1950-1955 then decreases (which makes sense as we know that the Cold War was occuring) If the data continued on I bet we would see the number of armed forces get lighter in color as the Vietnam War would then be starting. Overall I was was able to visually see GNP increase do to a few factors, which make sense as the more our population grows the more we produce and the occurance of war also influences an increase in production ( just more war time items ). Overall, I created a visual to help me with my questions and am now left wondering if the numebr of armed forces is double counted in the number of people also employed. 


